---
title: "RLQ Tutorial"
author: "Clement Garcia, Cefas"
date: "2025-04-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document provides an example data and R code for performing **RLQ analysis** (RLQ) and **Fourth Corner**, (4thC) RLQ combines three tables: site x species, site x environment and species x trait in one single ordination and 4thC tests the significance of different elements of the three tables.

For a complete working example, please refer to the [tutorial from Dray et al. (2016)](https://ndownloader.figshare.com/files/5628774) & [Combining the fourth-corner and the RLQ methods for assessing trait responses to environmental variation](https://doi.org/10.6084/m9.figshare.c.3306393.v1)

# 1.Preparation
## 1.1. Packages and data

he analyses require the R packages `ade4` and `tidyverse`.

```{r prep, echo=T, include = FALSE}
# Clear memory
rm(list=ls())
# Load packages
library(ade4) ## Perform RLQ & 4thC
library(tidyverse) ## data formatting and plotting (ggplot2)
```

If you get an error message, check that the R packages are installed correctly. If not, use the command: `install.packages(c("ade4", "tidyverse"))`.
The example dataset is available as the Rdata file `BenthicTrait_NL.Rdata`, available for download [here](https://github.com/Clem2012/MARBEFESWP3_VA/raw/refs/heads/main/processed_data/BenthicTrait_NL.RData).

```{r data}
load("C:/Users/cg05/OneDrive - CEFAS/Science/Project - EU/MARBEFES/WP3/CVA/MARBEFESWP3_VA/processed_data/BenthicTrait_NL.Rdata")
```
- `coord` stations coordinate (lat/lon)
- `fauna` Species abundance by site (the community matrix)
- `data.trait` Traits by species
- `lab.trait` Traits labels and code

```{r inspect}
# Inspect data tables 
dim(coord) 

dim(fauna)
names(fauna)

dim(data.trait)
names(data.trait)

unique(lab.trait$Trait)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
